\documentclass{sn-jnl}

\usepackage{graphicx}
\usepackage{amsmath}
\usepackage{amssymb}
\usepackage{booktabs}
\usepackage{multirow}
\usepackage{url}
\usepackage{hyperref}
\usepackage{tikz}
\usetikzlibrary{arrows.meta, positioning, shapes, fit}

\begin{document}

\title{From CPU-Orchestrated to GPU-Driven Storage: A Survey of GPU-Integrated I/O Architectures}

\author*[1]{\fnm{First} \sur{Author}}
\affil[1]{\orgname{Institution Name}, \orgaddress{\city{City}, \country{Country}}}

\abstract{While GPUs have become the dominant compute engines for modern data-intensive applications, including large-scale AI training and inference, a systematic understanding of their interaction with storage systems remains largely unexplored. This survey aims to bridge this gap by addressing the following key questions: (1) How have storage systems evolved to support GPU-centric workloads? (2) To what extent do existing architectures overcome the limitations of CPU-orchestrated I/O paths? (3) What are the emerging directions toward fully GPU-driven storage systems?

To answer these questions, we present a comprehensive survey of GPU-integrated I/O architectures. Specifically, we first characterize the architectural evolution from CPU-orchestrated to GPU-driven storage, and introduce a unified taxonomy spanning control plane ownership, data path realization, and deployment models, providing a structured lens to understand existing systems and frameworks.

Furthermore, we analyze key system design dimensions, including data movement, I/O scheduling, memory management, and fault tolerance, and distill cross-layer insights on bottleneck migration, metadata critical paths, queue-pair scalability, and memory registration overheads. Finally, we highlight emerging challenges and future directions, such as GPU-native I/O control, disaggregated storage for AI workloads, and composable memory--storage fabrics.

We hope this survey provides a principled understanding of the design space of GPU-integrated storage systems, offering both insights into their capabilities and limitations, and practical guidance for building next-generation GPU-driven data infrastructures.}

\keywords{GPU storage, I/O architecture, GPUDirect Storage, NVMe-oF, AI systems}

\maketitle

\input{sections/introduction}
\input{sections/related_work}
\input{sections/taxonomy}
\input{sections/figures}
\input{sections/design}
\input{sections/ai_workloads}
\input{sections/future}
\input{sections/conclusion}

\bibliography{ref}

\end{document}
